---
title: "Lab 3 - Reducing Crime"
author: "Clayton G. Leach, Karl I. Siil, Timothy S. Slade"
date: "July 23, 2018"
mainfont: 'Arial'
header-includes:
- \usepackage{caption}
- \usepackage{hhline}
- \usepackage{footnote}
- \usepackage{array}
- \usepackage{booktabs}
#- \usepackage{minted}
output: 
  pdf_document: 
    latex_engine: xelatex
    pandoc_args: "--pdf-engine-opt=-shell-escape"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(here) # TS: Added this to handle the user-agnostic run.
library(car)
```

# Introduction

Our client is running for office in the state of North Carolina (NC). Her campaign commissioned us to research the determinants of crime in NC to help her develop her platform regarding crime-related policy initiatives at the level of local government. This report explores a 1994 dataset from Cornwell & Trumball that provides county-level economic, demographic, and crime data. Our analysis describes the dataset, presents some initial summary statistics, develops three plausible models of the determinants of crime, and evaluates their accuracy and utility.

# Initial Exploratory Data Analysis (EDA)

```{r data_import, echo = FALSE, results = "hide"}
#here::dr_here() # TS: Use on first run to see why it fails on your machine, if it does.
here::here() # TS: Use when all set up, so we can mask the output.
#crime_raw <- read_csv('../crime_v2.csv', # TS: Commented out b/c using "here()" don't need ../
crime_raw <- read_csv('crime_v2.csv', col_types = cols( prbconv = col_double()))
problems(crime_raw) # TS: To comment out once we're ready to generate our draft report.
crime_raw[97, ]
```

## Missing Values

```{r drop_blank_rows}
# KS: Rows with no data
crime_na <- crime_raw %>% filter_all(any_vars(!is.na(.)))
# KS: Row with one back tick
crime_na %>% filter_all(any_vars(is.na(.))) %>% select(which(!is.na(.)))
crime_na <- crime_na %>% filter_all(all_vars(!is.na(.)))
```

Upon loading the data, we examine the `r nrow(crime_raw %>% filter_all(any_vars(is.na(.))))` rows that are missing data, finding that 5 are entirely blank and 1 contains only a backtick. We eliminate those to generate our working dataset.

```{r summary, include=FALSE}
crime_na %>% summary() # TS: I think we'll want to suppress this snippet in the draft we turn in...added include = FALSE
```

## Erroneous Duplicate Records

```{r find_dup_county}
crime_na %>% count(county) %>% filter(n > 1) # county 193 is an exact duplicate 
crime_na %>% filter(county == 193)
```

Continuing our QC, we note that one of the counties' records has been duplicated exactly. We therefore drop the duplicate record from our dataset.

```{r drop_dup_county}
crime_na <- crime_na %>% filter(!duplicated(.))
```

## Plausibility Checks for Variables

Three of our key variables of interest (`prbarr`, `prbconv`, and `prbpris`) represent probabilities and should therefore theoretically be in the range of 0:1.

```{r prob_beyond_range}
# look at weird 'probability' variables.
non_prob <- crime_na %>%
  filter(!between(prbarr, 0, 1) | !between(prbconv, 0, 1) | !between(prbpris, 0, 1))
```

Examining the data, we find `r nrow(non_prob)` counties have values for the "probability" variables that are outside of the expected range. In each case, it is either `prbconv` (10 records) or `prbarr` (1 record) that fall outside the range.

Per the notes accompanying our data, _The probability of conviction is proxied by the ratio of convictions to arrests..._ Given that definition, if not all suspects arrested are convicted, `prbconv` will be below 1. However, it may also exceed 1 if the number of exonerated suspects is exceeded by the number of suspects convicted of multiple charges. (See **[here](https://www.michigancriminallawyer-blog.com/2014/08/this-blog-is-based-upon.html)** for examples of multiple charges stemming from a single arrest.)

The notes on `prbarr` indicate _the probability of arrest is proxied by the ratio of arrests to offenses..._. If multiple suspects are arrested for a single offense, and this happens more frequently than offenses which do not lead to arrests, `prbarr` would indeed exceed 1.

In both cases, there are plausible explanations for the values we observe. Therefore we will not drop these records from our dataset. We will, however, subject them to further scrutiny.

```{r plausibility_checks, echo = FALSE, results = "hide", fig.show = "hide"}
plot(density(crime_na$year)) # TS: Confirmed all 1987
plot(density(crime_na$crmrte)) # TS: Confirmed all positive
plot(density(crime_na$prbarr)) # TS: Confirmed all positive
plot(density(crime_na$prbconv)) # TS: Confirmed all positive
plot(density(crime_na$prbpris)) # TS: Confirmed all positive
plot(density(crime_na$avgsen)) # TS: Confirmed all positive
plot(density(crime_na$polpc)) # TS: Confirmed all positive; in boxplot, 1 outlier
plot(density(crime_na$density)) # TS: Confirmed all positive; in boxplot, 8 outliers
plot(density(crime_na$taxpc)) # TS: Confirmed all positive; in boxplot, several outliers
plot(density(crime_na$west)) # TS: Confirmed all either 0 or 1
plot(density(crime_na$central)) # TS: Confirmed all either 0 or 1
plot(density(crime_na$urban)) # TS: Confirmed all either 0 or 1
plot(density(crime_na$pctmin80)) # TS: Confirmed all positive, between 0 and 100
plot(density(crime_na$wcon))
plot(density(crime_na$wtuc))
plot(density(crime_na$wtrd))
plot(density(crime_na$wfir))
plot(density(crime_na$wser)) # TS: Some serious outliers at > $2000
plot(density(crime_na$wmfg))
plot(density(crime_na$wsta))
plot(density(crime_na$wloc))
plot(density(crime_na$mix))
```

Examining the remainder of our data, we found no substantial evidence of _top-coded_ or _bottom-coded_ (i.e., truncated) variables which might bias our regression models. However, there is an extreme outlier in `wser`, the variable indicating the county's weekly wage in the service industry.

# Research Question and Model-Building

Our **research question** is the following: _What are the determinants of crime at the county level?_

We face a key limitation: our data does not give us visibility into crime, it only gives us insight into the official _crime rate_. The crime rate is a function not only of crimes committed but also of various factors, some of which may be unobservable. For instance, poor community-police relations may bias crime rates downward if an area's residents **[do not report all the crimes they observe or experience](https://www.washingtonpost.com/news/acts-of-faith/wp/2018/04/19/churches-make-a-drastic-pledge-in-the-name-of-social-justice-to-stop-calling-the-police/?utm_term=.ac2ea8ee1523)**. Converseley, those poor relations may also bias crime rates upward if police officers engage in **[predatory policing practices](https://www.huffingtonpost.com/2015/05/04/st-louis-county-predatory-policing_n_7201964.html)** and the community lacks the wherewithal to fight back. As we report our findings we will make note of potential bias that results from our inability to observe and analyze critical variables.

## Explanatory variables of interest

The table below details several main variables of interest we will use to build and refine our model.

\begin{center} % https://community.rstudio.com/t/inserting-table-in-r-markdown/7260/4
\captionof{table}{Hypothesized Primary Determinants of Observed Crime Rate}
\begin{tabular}{|l|l|p{7cm}|p{2cm}|}
\hline
Variable Name & Explanation & Reasoning & Transformation Applied \\ \hhline{|=|=|=|=|}
\texttt{polpc} & \textit{police per capita} & Police may act as a deterrent to crime, may increase the observed crime rate, or both. & <none> \\
\texttt{pctymle} & \textit{percent young male} & Young males commit and are charged with a disproportionate share of crimes & <none> \\
\texttt{density} & \textit{people per sq. mile} & Greater population density increases opportunity for crimes to be committed and reported & <none> \\
\texttt{taxpc} & \textit{tax revenue per capita} & Lower tax revenues may be associated with poorer community-government relations, greater economic hardship, and less policing \footnote{This may introduce collinearity with several other variables} & $log_{10}$  \\
\texttt{prbarr} & \textit{`probability' of arrest} & Greater probability of arrest may serve a deterrent function & <none> \\
\texttt{prbconv} & \textit{`probability' of conviction} & Greater probability of conviction may serve a deterrent function & <none> \\
\texttt{prbpris} & \textit{`probability' of prison sentence} & Greater probability of sentencing may serve a deterrent function & <none> \\
\texttt{avgsen} & \textit{average sentence, in days} & Harsher sentencing practices may serve a deterrent function & <none> \\
\texttt{pctmin80} & \textit{percent minority in 1980} & Minorities are disproportionately arrested and convicted of crimes & <none> \\
\hline
\end{tabular}
\end{center}

\[
%\begin{table}
%\centering
%\begin{tabular}{l>{\raggedright\arraybackslash$}p{2cm}<{$}{\raggedright\arraybackslash$}p{6cm}<{$}c}
%\toprule
%Variable Name & Explanation & Reasoning & Transformation Applied \\ \midrule
%\texttt{polpc} & \textit{police per capita} & Police may act as a deterrent to crime, may increase the observed crime rate, or both. & <none> \\
%\texttt{pctymle} & \textit{percent young male} & Young males commit and are charged with a disproportionate share of crimes & <none> \\
%\texttt{density} & \textit{people per sq. mile} & Greater population density increases opportunity for crimes to be committed and reported & <none> \\
%\texttt{taxpc} & \textit{tax revenue per capita} & Lower tax revenues may be associated with poorer community-government relations, greater economic hardship, and less policing \footnote{This may introduce collinearity with several other variables} & $log_10$  \\
%\texttt{prbarr} & \textit{`probability' of arrest} & Greater probability of arrest may serve a deterrent function & <none> \\
%\texttt{prbconv} & \textit{`probability' of conviction} & Greater probability of conviction may serve a deterrent function & <none> \\
%\texttt{prbpris} & \textit{`probability' of prison sentence} & Greater probability of sentencing may serve a deterrent function & <none> \\
%\texttt{avgsen} & \textit{average sentence, in days} & Harsher sentencing practices may serve a deterrent function & <none> \\
%\texttt{pctmin80} & \textit{percent minority in 1980} & Minorities are disproportionately arrested and convicted of crimes & <none> \\
%\midrule
%\end{tabular}
%\caption{Hypothesized Primary Determinants of Observed Crime Rate}
%\label{pntable}
%\end{table}
\]



```{r lm_1}
lm_mod1a <- with(crime_na, lm(crmrte ~ polpc + pctymle))
lm_mod1b <- with(crime_na, lm(crmrte ~ polpc + pctymle + density))
lm_mod1c <- with(crime_na, lm(crmrte ~ polpc + pctymle + density + log(taxpc, base = 10)))
lm_mod1d <- with(crime_na, lm(crmrte ~ polpc + pctymle + density + log(taxpc, base = 10) +
                                  prbarr + prbconv + prbpris))
lm_mod1e <- with(crime_na, lm(crmrte ~ polpc + pctymle + density + log(taxpc, base = 10) +
                                  prbarr + prbconv + prbpris + avgsen))
#lm_mod1 <- with(crime_na, lm(crmrte ~ polpc + pctymle + density + log(taxpc, base = 10) + prbarr + prbconv + prbpris + avgsen))
```

Once \texttt{pctymle} and \texttt{density} are included in the model, \texttt{polpc} loses its significance.



