---
title: "Lab 03 EDA"
author: "Timothy S. Slade"
date: "July 12, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, tidy = TRUE)
tidy.opts=list(blank=FALSE, width.cutoff=60)
```

## Workspace Preparation

In the following chunk we load the libraries we intend to use in this EDA.

```{r libraries}
library("here")
library("dplyr")
library("ggplot2")
library("car")
```

Now we set the workspace and load our dataset.

```{r workspace}
here::dr_here()
df <- data.frame(read.csv(file = "crime_v2.csv"))
```

The dataset contains `r length(names(df))` variables and `r nrow(df)` observations.

```{r characterizing_df}
names(df)
summary(df)
```

We have the following guidance from the instructors regarding the contents of our initial Lab03 submission:

1. State your research question clearly, succinctly.
2. State your goals precisely, they need to be in finite terms, not open ended.
3. Explain clearly what motivates your research, your interest in this dataset and wrap around it how you plan to carry this research out(clear steps, iterative approach)
4. Do #1,2 first, determine your motivation first and in the end stick to 2-3 variables that most interest you, why, why you decided to pick them vs not some others
5. Be sure that your chosen variables are directly related to your proposed questions.
And finally Eric added that do read and understand the units 9,10,11 carefully as that will help understand key elements of this approach. 

# Draft Report

_Your task is to examine the data to help the campaign understand the determinants of crime and to generate policy suggestions that are applicable to local government._

Crime rates and police activity are likely to exhibit regional effects.